beautier: BEAUti for R

### 2017-12-19 article.tex
beastscriptr is beautier

% \newcommand*\beastscriptr{\verb;beastscriptr;}

\title{beastscriptr: BEAUti for R}
\author{Rich\`el J.C. Bilderbeek}

@@

\begin{abstract}
  \textbf{1. }
  Here, I present a package, \verb;beastscriptr;, for the R programming language. \\
  \textbf{2. }
    \verb;beastscriptr; allows for scripted use of the BEAST2 phylogenetics tool, 
    by creating BEAST2 input files from an R function call. \\
  \textbf{3. }
    I describe \verb;beastscriptr; usage, the novel functionality it provides
    compared to BEAUti, and give some minimal examples. \\
  \textbf{4. }
    As \verb;beastscriptr; is free, libre, open-source and designed to be extended, 
    I conclude by describing the current development of the package \\
\end{abstract}

@@

BEAST2 \cite{bouckaert2014beast} is a popular Bayesian phylogenetics tool.
BEAST2 is bundled with the program called BEAUti \cite{drummond2012bayesian},
which  allows for a user-friendly way to create the
input files that BEAST2 needs. BEAUti is convenient to get started, yet
does not allow for the scripted use of BEAST2. For creating BEAST2 input
files from a script, beastscriptr is the R package facilitating this:
it creates a BEAST2 input files from an R function call.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=6cm, semithick]   
  \tikzstyle{every state}=[]
  \node[state] (A) [rectangle] {
    \includegraphics[width=0.2\textwidth]{alignment.png}
    \includegraphics[width=0.2\textwidth]{thought_cloud.png}
  };   
  \node[state] (E) [below of=A, rectangle] {
    \includegraphics[height=0.1\textheight]{xml.png}
  };   
  \node[state] (F) [rectangle, below of=E] {\includegraphics[width=0.3\textwidth]{DensiTreeExample2.jpg}};
  \path (A) edge [anchor = east] node {\includegraphics[height=0.15\textheight]{beastscriptr_logo.png}} (E)
        (A) edge [anchor = west] node {\includegraphics[height=0.15\textheight]{BeautiSiteModel.png}} (E)
        (E) edge [] node {\includegraphics[width=0.2\textwidth]{Beast2LogoRat.png}} (F); 
  \end{tikzpicture}

@@

The goal of the R package \verb;beastscriptr; is to create the 
same BEAST2 input files that BEAUti creates, but from one R function call 
instead.

By creating BEAST2 input files in BEAUti, 
the desired output of beastscriptr was determined. 
The functionality of BEAUti is added gradually (and is still growing).

\verb;beastscriptr; has only parts of the functionality of BEAUti, yet
can be extended easily. beastcriptr, on the other hand, does allow for specifying
to use phylogenies of fixed crown age.

\verb;beastscriptr; has a novel functionality not built into BEAUti yet:
it allows for using phylogenies of a fixed crown age.

@@

As in BEAUti, to create a BEAST2 input file, 
one needs an alignment and model settings.
With those settings at their default, \verb;beastscriptr; is easy to use:
Create a BEAST2 input file with name \verb;beast2.xml; from 
a FASTA file with name \verb;alignment.fas; is done like this:

\begin{lstlisting}[language=R, caption=Simplest example]
library(beastscriptr)
create_beast2_input_file(
  "alignment.fas",
  "beast2.xml"
)
\end{lstlisting}

@@

Novel about beastscriptr is that it allows for specifying a fixed crown age:

\begin{lstlisting}[language=R, caption=Example with fixed crown age]
create_beast2_input_file(
  "alignment.fas",
  "beast2.xml"
  fixed_crown_age = TRUE,
  initial_phylogeny = beastscriptr::fasta_to_phylo(
    fasta_filename = "alignment.fas", 
    crown_age = 15
  )
)
\end{lstlisting}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{beastscriptr development and other resources}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

beastscriptr successfully creates valid BEAST2 input files. Unlike BEAUti,
beascriptr allows for setting a fixed crown age.

beastscriptr does not support the full functionality of BEAUti. Considering
the size, age and number of plugins, this would be close to impossible.
To compensate for this, an extensible software architecture is used.

beastscriptr has minimal support for calling BEAST2 from within R and does
so for testing purposes.

@@

The package is free and is available from the official R package archive at 
http://cran.r-project.org/src/contrib/PACKAGES.html\#beastscriptr. 
beastscriptr is licensed under the GNU General Public License.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Citation of beastscriptr}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Scientists using beastscriptr in a published paper should cite this
article. Users can additionally cite the beastscriptr package 
directly. Citation information can be obtained by typing:

\begin{lstlisting}[language=R]
> citation("beastscriptr")
\end{lstlisting}

>>>>

% \newcommand*\beautier{\verb;beautier;}

\title{beautier: BEAUti for R}
\author{Rich\`el J.C. Bilderbeek}

@@

\begin{abstract}
  \textbf{1. }
  Here, I present a package, \verb;beautier;, for the R programming language. \\
  \textbf{2. }
    \verb;beautier; allows for scripted use of the BEAST2 phylogenetics tool, 
    by creating BEAST2 input files from an R function call. \\
  \textbf{3. }
    I describe \verb;beautier; usage, the novel functionality it provides
    compared to BEAUti, and give some minimal examples. \\
  \textbf{4. }
    As \verb;beautier; is free, libre, open-source and designed to be extended, 
    I conclude by describing the current development of the package \\
\end{abstract}

@@

BEAST2 \cite{bouckaert2014beast} is a popular Bayesian phylogenetics tool.
BEAST2 is bundled with the program called BEAUti \cite{drummond2012bayesian},
which  allows for a user-friendly way to create the
input files that BEAST2 needs. BEAUti is convenient to get started, yet
does not allow for the scripted use of BEAST2. For creating BEAST2 input
files from a script, beautier is the R package facilitating this:
it creates a BEAST2 input files from an R function call.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
  \centering
  \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=6cm, semithick]   
  \tikzstyle{every state}=[]
  \node[state] (A) [rectangle] {
    \includegraphics[width=0.2\textwidth]{alignment.png}
    \includegraphics[width=0.2\textwidth]{thought_cloud.png}
  };   
  \node[state] (E) [below of=A, rectangle] {
    \includegraphics[height=0.1\textheight]{xml.png}
  };   
  \node[state] (F) [rectangle, below of=E] {\includegraphics[width=0.3\textwidth]{DensiTreeExample2.jpg}};
  \path (A) edge [anchor = east] node {\includegraphics[height=0.15\textheight]{beautier_logo.png}} (E)
        (A) edge [anchor = west] node {\includegraphics[height=0.15\textheight]{BeautiSiteModel.png}} (E)
        (E) edge [] node {\includegraphics[width=0.2\textwidth]{Beast2LogoRat.png}} (F); 
  \end{tikzpicture}

@@

The goal of the R package \verb;beautier; is to create the 
same BEAST2 input files that BEAUti creates, but from one R function call 
instead.

By creating BEAST2 input files in BEAUti, 
the desired output of beautier was determined. 
The functionality of BEAUti is added gradually (and is still growing).

\verb;beautier; has only parts of the functionality of BEAUti, yet
can be extended easily. beastcriptr, on the other hand, does allow for specifying
to use phylogenies of fixed crown age.

\verb;beautier; has a novel functionality not built into BEAUti yet:
it allows for using phylogenies of a fixed crown age.

@@

As in BEAUti, to create a BEAST2 input file, 
one needs an alignment and model settings.
With those settings at their default, \verb;beautier; is easy to use:
Create a BEAST2 input file with name \verb;beast2.xml; from 
a FASTA file with name \verb;alignment.fas; is done like this:

\begin{lstlisting}[language=R, caption=Simplest example]
library(beautier)
create_beast2_input_file(
  "alignment.fas",
  "beast2.xml"
)
\end{lstlisting}

@@

Novel about beautier is that it allows for specifying a fixed crown age:

\begin{lstlisting}[language=R, caption=Example with fixed crown age]
create_beast2_input_file(
  "alignment.fas",
  "beast2.xml"
  fixed_crown_age = TRUE,
  initial_phylogeny = beautier::fasta_to_phylo(
    fasta_filename = "alignment.fas", 
    crown_age = 15
  )
)
\end{lstlisting}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{beautier development and other resources}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

beautier successfully creates valid BEAST2 input files. Unlike BEAUti,
beascriptr allows for setting a fixed crown age.

beautier does not support the full functionality of BEAUti. Considering
the size, age and number of plugins, this would be close to impossible.
To compensate for this, an extensible software architecture is used.

beautier has minimal support for calling BEAST2 from within R and does
so for testing purposes.

@@

The package is free and is available from the official R package archive at 
http://cran.r-project.org/src/contrib/PACKAGES.html\#beautier. 
beautier is licensed under the GNU General Public License.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Citation of beautier}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Scientists using beautier in a published paper should cite this
article. Users can additionally cite the beautier package 
directly. Citation information can be obtained by typing:

\begin{lstlisting}[language=R]
> citation("beautier")
\end{lstlisting}
